---
title: "DSO 499: Essentials of Business Data Analysis Using R"
author: "Intro to ggplot2"
date: "Fall 2019"
output:
  
  pdf_document: default
  number_sections: true
  theme: united
  toc: yes
  toc_depth: 3
header-includes:
   - \usepackage{color}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,eval = F, width = 50, tidy.opts=list(width.cutoff=40), tidy=TRUE)
```


# Scatterplots 

@. Load the `ggplot2` package into R. 

```{r, warning=FALSE}
#load ggplot2 so that we can use the function and data that comes with it
# you'll have to comment the install.packages() function once you've installed it on your machine.

# install.packages("ggplot2")
library(ggplot2) 

```

@. Use the `mpg` dataset that comes with the `ggplot2` package, and create a scatterplot that shows the relationship between the engine displacement `displ` and the highway miles per gallon variable `hwy`.
```{r}
?mpg
?ggplot

ggplot(mpg, aes(x=displ, y=hwy)) + 
  geom_point()


```


@. Use color to distiguish between different types (`class`) of cars in the preivous plot. 

```{r}
?mpg
?ggplot

ggplot(mpg, aes(x=displ, y=hwy, size = 2)) + 
  geom_point(aes(col=class))



```



@. In the previous plot, we used assigned a categorical variable to the `color` aesthetics. In this exercises, try to add the `color` aesthetics to the `geom_point()` line of code. Explain what happened. 

```{r}
?mpg
?ggplot

ggplot(mpg, aes(x=displ, y=hwy, color = class)) + 
  geom_point(color ="red")



```

@. Use the `size` aesthetic to show the different types of cars (`class`) instead of the color aesthetic. 

```{r}
?mpg
?ggplot

ggplot(mpg, aes(x=displ, y=hwy, size = class)) + 
  geom_point()

ggplot(mpg, aes(x=displ, y=hwy, size = cyl)) + 
  geom_point(aes(col = class))

```

### Facets

@. Use the function `facet_wrap()` to show the relationship between `displ` and `hwy` for each of the car types (`class`) in a seperate graph. 

```{r}
?mpg
?ggplot

ggplot(mpg, aes(x=displ, y=hwy)) + 
  geom_point() +
  facet_wrap(~class)

ggplot(mpg, aes(x=displ, y=hwy)) + 
  geom_point() +
  facet_wrap(~class, nrow = 2)

ggplot(mpg, aes(x=displ, y=hwy)) + 
  geom_point() +
  facet_wrap(cyl~class)

```

# Bar Graphs 

@. Use the `str()` function to explore the `diamonds` dataset that comes with `ggplot2` package. 

```{r}
?diamonds
str(diamonds)
```

@. Create a barchart that shows the count distribution for each of the `cut` levels of the diamonds. 

```{r}
?diamonds
?geom_bar

ggplot(diamonds, aes(x = cut)) +
  geom_bar()

ggplot(diamonds, aes(x = cut)) +
  geom_bar(fill = "purple")

ggplot(diamonds, aes(x = cut)) +
  geom_bar(aes(col = "purple"))
```

@. Check what happens if you try to create a count bar chart for the variable `carat`. 

```{r}
?diamonds
?geom_bar

# Note that 'carat' BAR CHARTS don't look good
ggplot(diamonds, aes(x = carat)) +
  geom_bar()

# Let's make BOXPLOTS instead

ggplot(diamonds, aes(x = cut)) +
  geom_boxplot(aes(x=cut, y= carat))

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_boxplot(aes(group = cut_width(carat, 0.5)))


# Now, let's make HISTOGRAMS instead
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 10) +
  ylab("Number of Diamonds") +
  xlab("U.S. Dollars") +
  labs(title = "Price Distribution")

```

@. Load the `gcookbook` pacakge to R. 

```{r}
#install.packages("gcookbook")
library(gcookbook)
```

@. Investigate the contents of the `BOD` dataset that comes with `gcookbook` R package. 

```{r}
?BOD
str(BOD)
```

@. Create a bar chart that shows you the biochemical oxygen demand for each time in an evaluation of water quality.

```{r}
?BOD
str(BOD)

ggplot(BOD, aes(x = Time, y = demand)) +
  geom_col() +

```

@. How can you improve the previous graph?

```{r}
?BOD
str(BOD)

ggplot(BOD, aes(x = as.factor(Time), y = demand)) +
  geom_col()

```

@. Change the with of each bar in the previous graph into 0.5. 

```{r}
?BOD
str(BOD)

ggplot(BOD, aes(x = as.factor(Time), y = demand)) +
  geom_col(width = 0.5)

```
@. Change the color of the bars in the previous graph into `lightblue`.

```{r}
?BOD
str(BOD)

ggplot(BOD, aes(x = as.factor(Time), y = demand)) +
  geom_col(width = 0.5, fill = "lightblue", col = "red")

```

### Adding Labels to a Bar Graph

To do so, we will add a new layer to our plot using `geom_text()`.


@. For the BOD barchart created before, add the oxygen demand value on the bars for each of the Times. 
```{r}
?BOD
str(BOD)

ggplot(BOD, aes(x = as.factor(Time), y = demand)) +
  geom_col(width = 0.5, fill = "lightblue", col = "red") +
  geom_text(aes(label = demand, vjust = -0.5))

```

### Horizontal Bar Charts

@. Create a horizontal barchart version of the BOD barchart using the `coord_flip()` function. 

```{r}
?BOD
str(BOD)

ggplot(BOD, aes(x = as.factor(Time), y = demand)) +
  geom_col(width = 0.5, fill = "lightblue", col = "red") +
  coord_flip()

```

### Grouping Bars Together

@. Create a bargraph that shows how each `cut` level in the diamonds dataset is distribued among the different `clarity` levels for the diamonds.

```{r}
ggplot(diamonds, aes(x = cut, fill = clarity)) +
geom_bar(position = "dodge", col = "black")
```

### Using Colors in a Bar Graph

@. Exlpore the `uspopchange` dataset that comes with the `gcookbook` R package. 

```{r}
str(uspopchange)
```


@. Create a barchat that shows the % change in population for the top 10 states. The bars should be colored based on the region of each state (default colors).



@. Use the `scale_fill_manual()` function to update the colors of the previous chart as follows: South (`pink`), and West (`lightblue`). 


@. Reorder the bars in the previous barchart to decreasing order. 


### Coloring Negative and Positive Bars Differently

@. Explore the `climate` dataset from the `gcookbook` R package. 

```{r}
str(climate)
```

@. Create a dataframe for observations (Year 1900 and after) in the climate dataset that have Berkeley as their datasource. Save it in a variable called  `subclimate`.


@. Create a new `boolean` variable `Pos` in the `sublclimate` dataset. This variable is `T` if `Anomaly10y` is positive otherwise, it will be `F`. 


@. Use the `subclimate` dataset to create a barchart that shows the distribution of the `Anomaly10y` values over the years. Use coloring to distinguish between postive and negative anomalies. 


@. Update the colors in the previous graph to `lightblue` for positive anomalies, and `red` for negative anomalies. 

# Cleveland Dot Plot or Dot Plots

@. Explore the `tophitters2001` dataset that comes in the `gcookbook` R package.


@. Create a dotplot that shows the name as well as the average batting score for the top 10 players. 



